## Reproducible Research Assignment 1
### Loading and preprocessing data
1. Download data from https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip 
2. Unzip the file using standard platform tooling
3. load data

```{r load_data}
activityData <- read.csv("activity.csv")
```

### Information on total number of steps taken per day
Plot of frequency of total number of steps taken during a day
```{r StepsDays}
summarySteps <- aggregate(activityData$steps ~activityData$date,data = activityData, sum)
meanTotalSteps <- mean(summarySteps[,2])
medianTotalSteps <- median(summarySteps[,2])
hist(summarySteps[,2],main="Total Number of Steps Taken during a Day",xlab="Number of Steps",ylab="Number of Days")
```
mean number of steps per day	`r meanTotalSteps`

median number of steps per day 	`r medianTotalSteps`

### Information on average daily activity pattern 
Plot of number of steps taken on average during each five minute interval
```{r StepsHours}
summaryIntervals <- aggregate(activityData$steps ~activityData$interval,data = activityData, sum)
names(summaryIntervals) <- rbind("Intervals","Steps")
plot(summaryIntervals,type='l',main = "Average Number of Steps During 24 hours",xlab="5 minute interval from Midnight")
maxSteps <- max(summaryIntervals$Steps)
```
Maximum average number of steps taken during five minute interval `r maxSteps`

### Impute missing values
Use the mean value of the five minute time interval if there is no value in the data set for the number of steps

```{r replaceNA}
numberNA <- sum(is.na(activityData[,1]))
listNA <- is.na(activityData[,1])
updatedActivityData <- activityData
numberIntervals <- dim(activityData)[1]
for (n in 1:numberIntervals)
	{
	if (listNA[n])
		{
		updatedActivityData[n,1] <-  summaryIntervals[(updatedActivityData[n,3] == summaryIntervals[,1]),2]
		}
	}
```

Plot of frequency of total number of steps taken during a day having imputed any missing number of steps

```{r stepsDaysImputed}
summaryStepsImputed <- aggregate(updatedActivityData$steps ~ updatedActivityData$date,data = updatedActivityData, sum)
meanTotalStepsImputed <- mean(summaryStepsImputed[,2])
medianTotalStepsImputed <- median(summaryStepsImputed[,2])
hist(summarySteps[,2],main="Total Number of Steps Taken during a Day",xlab="Number of Steps",ylab="Number of Days")
diffmean <- meanTotalStepsImputed - meanTotalSteps
diffmedian <- medianTotalStepsImputed - medianTotalSteps
```
mean number of steps per day using imputed data		`r meanTotalStepsImputed`

median number of steps per day using imputed data 	`r medianTotalStepsImputed`

mean difference between raw data and data with no missing values 	`r diffmean`

median difference between raw data and data with no missing vlaues 	`r diffmedian`

### Average steps for interval during week days versus week ends
Add weekday or weekend to imputed data
```{r createDayofWeel}
week_days <- weekdays(as.Date(updatedActivityData$date,format ="%Y-%m-%d"))
weekdayActivityData <- cbind(updatedActivityData,week_days)
weekend <- (week_days == "Saturday" | week_days == "Sunday")
weekend_factor <- factor(weekend,labels = c("Weekday","Weekend"))
weekdayActivityData <- cbind(weekdayActivityData,weekend_factor)
weekend_steps <- aggregate(weekdayActivityData$steps ~ weekdayActivityData$interval + weekdayActivityData$weekend_factor,data = weekdayActivityData, sum)
```
Plot results
```{r weekendPlot}
names(weekend_steps) <- c("intervals","time","average_steps")
q <-qplot(intervals,average_steps,data=weekend_steps,geom="line",colour=time)
print(q)
```